---
title: "GitHub Workflow"
output: html_document
date: "2024-06-04"
---

These steps are talked about in more detail in the book:

- https://datatrail-jhu.github.io/DataTrail/creating-a-repository.html
- https://datatrail-jhu.github.io/DataTrail/cloning-a-repository.html
- https://datatrail-jhu.github.io/DataTrail/pushing-and-pulling-changes.html

## Steps for linking GitHub to Posit Cloud

Github is a really powerful opensource website that works with git, a version control system.

To use it in posit cloud we need to set up credentials with our github account.

Let's use the `usethis` package to help us link to our GitHub account.

Only need to do this once per project.

```{r}
install.packages("usethis")
library("usethis")
```

We need to set up credentials.

We'll create a Personal Access Token
```
usethis::create_github_token()
```

We'll give posit cloud this token
```
gitcreds::gitcreds_set()
```

We'll also give our user name and email.
```
use_git_config(user.name = "cansavvy", user.email = "cansav09@gmail.com")
```

We can run this function to double check we've set this up.

```{r}
git_sitrep()
```


## Rundown on GitHub vocab

Repository = house for your project files
Cloning = Making a copy of the repository for posit cloud (or wherever you are working)
Personal Access Token = A key or fancy password to give us access to GitHub
Commit = Adding some file changes to a branch
Pushing = A transfer of files from where we are working to GitHub

## Creating a pull request

Create a new branch to work from. (that's different than the default main branch).

```{r}
pr_init(branch = "new-branch")
```

Make a file change (and save it).

In the Git tab, check the "staged" box next to the file change you made.

Click the commit button.
Write a descriptive commit message.
Click "commit". Now we can close this window.

```{r}
pr_push()
```

This will open up a window in Github that creates a pull request.

Pull Request = Its a method to suggest changes to the files
