---
title: "Exercise Intro to R"
author: "Candace Savonen"
date: '2022-08-09'
output: html_document
---

## Exercise 1: Installing packages from GitHub

1. Install a package called 'devtools'. 

```{r}
install.packages("devtools")
```

2. Attach the library of 'devtools'.

```{r}
library(devtools)
```

3. Practice installing a package from github using this command: 

This package can be seen here on GitHub: https://github.com/psolymos/KnockKnockJokes 

```{r}
install_github("psolymos/KnockKnockJokes")
```

4. You can use this package to tell you knock knock jokes by doing (remember to attach the library first). 

```{r}
library(KnockKnockJokes)
KnockKnock()
```

## Exercise 2: Make a data frame from scratch.

1. Make a vector that has 5 items in it using this function: `c()`

```{r}
animals <- c("dog", "cat", "mouse", "elephant", "cheetah")
```

2. If you don't know how to use this function look up the function with `?c`

3. Now make 2 more vectors that are also 5 items long in the same way. 

```{r}
fur_colors <- c("brown", "orange", "gray", "dark gray", "spotted")
cuddliness <- c(5, 3, 3, 2, 0)
```

4. Take all 3 vectors and put them together in one data frame using this function:

```{r}
animals_df <- data.frame(animals, fur_colors, cuddliness)
```

5. Look at one of the columns using a `$` to pull out one of the variables you used to make this data frame.

```{r}
animals_df$animals
```

6. Use the `readr::write_csv()` function to write your data frame to a CSV file. Make sure that you specify the file name using the file argument. Look up documentation about write_csv using ? for more details. Remember to put the filename you specify needs to be put in quotes.

```{r}
dir.create("data")
readr::write_csv(animals_df, file = "data/animals.csv")
```

## Exercise 3: Read in data from googlesheets and make a basic plot

1. Take a look at these movies in this [googlesheet](https://docs.google.com/spreadsheets/d/1okTb4MTllkDWlj0daHNfA8RwOPhLQkWtpQN_znJJJmk/htmlview?hl=en). 

2. Use the `googlesheets4` package and the `read_sheet` function to read in one of the sheets here. You can use the sheets argument to specify which sheet and genre you'd like to read in. 

```{r}
install.packages("googlesheets4")
library(googlesheets4)

comedies <- read_sheet("https://docs.google.com/spreadsheets/d/1okTb4MTllkDWlj0daHNfA8RwOPhLQkWtpQN_znJJJmk/htmlview?hl=en", sheet = 2)
```

3. Using the `max()` function, find whats the most number of oscars won by a film in the genre you read in. 

```{r}
max(comedies$`Oscars won`)
```

4. Rename the Oscars column so it is called `oscars_won`. 

```{r}
library(dplyr)
comedies <- rename(comedies, oscars_won = `Oscars won`)
```
5. Using the `dplyr::filter()` function filter to only movies that won Oscars. (Get rid of movies that have `NA` for Oscars) 

This will look like this:

```{r}
dplyr::filter(comedies, !is.na(oscars_won))
```

6. Assign this data frame as a new object. 

```{r}
oscar_winning_comedies <- dplyr::filter(comedies, !is.na(oscars_won))
```

7. Use the plot() function to plot the number of oscars by movie. Where `x = Film` and `y = oscars_won`. 

```{r}
plot(x = as.factor(oscar_winning_comedies$Film), 
     y = as.numeric(oscar_winning_comedies$oscars_won), 
     xlim = c(0,5), 
     las = 2)
```


8. Save this plot to a PNG using the Export button. 
