---
title: "File Organization Exercise"
author: "Candace"
date: '2022-08-04'
output: html_document
---

This exercise corresponds to the [File Organization course in DataTrail](https://datatrail-jhu.github.io/03_fileorganization/setting-up-data-science-project-folders.html). 

Note how I just made a link above. I used brackets to put a name for the link, followed by parantheses with the link that I want inside there. 

In my RMarkdown I can have notes in between my code. 

To make code chunks I can type the below by hand: 

```{r}
```

Or I can use +chunk button on the RStudio menu. 

## Install packages 

I should install whatever packages my R Markdown needs first. 

For this one, we only need `readr`. So here's how I install it. 

```{r}
install.packages("readr")
```

To use this package I need to use `library()` first though. 

```{r}
library(readr) 
```

In my files pane, I need to have the following folders: 
- `code` 
- `data`
- `figures`
- `products`

To check that these folders exist, I am going to use the `dir()` function. 

```{r}
dir()
```


Hopefully the names of the folders I listed above printed out here. 

If I want to know whats in a folder. I can use this `dir()` function again but put the name of a folder **in quotes**.

```{r}
# What is in the data folder? This code line below will print that out
dir("data")
```

The file I want to read in is called "git_project.csv" and its located in my "data" folder. 

Let me try to read that in:

```{r}
# Reading in a CSV (but I'm not assigning it as anything yet)
read_csv("data/git_project.csv")
```

Now I'm going to read this in again but I'm going to assign it (by using `<-`) to be called `music_data`. 

```{r}
music_data <- read_csv("data/git_project.csv")
```

Now I will see it show up in my environment panel (in the upper right corner). 

I can also use `View()` to have it pop up in a window so I can see these data. (but I can put that in this R Markdown document). I can type `View(music_data)` in the Console window though. 

If I want images to show up in my notebook I can do a similar thing that I did for links. I can use brackets followed by parantheses. But I have to point to where I have the image file. 

In this instance, I have an image file called `chubby_checker.jpg` that's located in the `figures` folder. So I can do the below thing:

[Album image](figures/chubby_checker.jpg)


Now to render the pretty version of this RMarkdown document, I first want to save this document (Click Command + S for a shortcut). 
Then I can click the Knit button above. (It has a yarn and knitting needle image next to it). 

Note that if there are any problems with my code that my html will not print out and I'll get an error. So I'll need to fix that first. 

Then it will print out a nice html. If I click that html I will see the pretty version of this document. 

