---
title: "Data Tidying Exercise"
output: html_document
date: "2022-08-17"
---

## Exercise 1: Saving and reading files

Saving to CSV or TSV files is generally fine, but sometimes you might have a R object that doesn't fit into a table format and you'd like to save it *exactly* how it is saved in R. 

For this purpose, RDS files are a great option. If you save an R object in an RDS file, then you can read it back into your R environment and it will be exactly how you saved it. 

To practice with RDS files, we can use an already existing object in R, `iris`. This dataset exists in R for practice purposes. Let's take a look at it. 

```{r}
# Print out iris to see what it looks like. 
iris
```

Let's fix some things in this dataset before we save it to RDS. 

### Column renaming 

Let's change column names to be in a more standard format using `dplyr::rename`. Instead of having the column names like `Sepal.Length`, change them to be like this: `sepal_length`. Where you've replace the `.` with an underscore `_` and turn capitalized letters into all lower. 

Save this new version of the dataset as `iris_cleaned`. 

```{r}
iris_cleaned <- 
```

Use the `saveRDS()` function to save `iris_cleaned` object to a file called "iris_cleaned.rds" that is located in the `data` folder.

```{r}

```

## Read in an RDS file

Now that you have an RDS file, let's see what it looks like when you read the "data/iris_cleaned.rds" file in using the `readRDS()` function. Assign this re-read object the name `iris_cleaned_reread`.

```{r}
iris_cleaned_reread <- 
```

Use the function `all.equal()` to see if `iris_cleaned_reread` (the read in version) and `iris_cleaned` (the original version) are indeed the same. 

```{r}

```

If you got a `TRUE` that means that they are identical R objects. 

## Exercise 2: Tidying NBA Finals Data

In this exercise we will work with [NBA final teams stats](https://www.kaggle.com/daverosenman/nba-finals-team-stats) on Kaggle. 
There are two .csv files. 
The 'champs.csv' file contains game-by-game team totals for the championship team from every finals game between 1980 and 2017. 
The 'runnerups.csv' contains game-by-game team totals for the runner-up team from every finals game between 1980 and 2017. The 1980 NBA Finals was the first Finals series since the NBA added the three point line.


### Download the data

Go to https://www.kaggle.com/daverosenman/nba-finals-team-stats to get the data 

Click the Download button in the upper right corner. 

In order to download data on Kaggle, you'll need to create a Kaggle account.
You can do so by linking your Google account. 

Unzip these data after you've downloaded them to your chromebook. 

### Make a data folder 

```{r}
# Use this command to make a new folder
dir.create("data", showWarnings = FALSE)
```

Use the Upload button in RStudio in the bottom right panel to upload these data RStudio cloud and specifically to your `data` folder.

To check if you've successfully uploaded these data in the right place, you can run this command. It will tell you `TRUE` if the `championsdata.csv` file is in the right place. 

```{r}
# Check for championsdata.csv in the data folder
file.exists(file.path("data", "championsdata.csv"))
```

We can do the same command for the `runnersupdata.csv` file.

```{r}
# Check for championsdata.csv in the data folder
file.exists(file.path("data", "runnerupsdata.csv"))
```

### Read in the data 

Use the readr::read_csv() function to read in the `championsdata.csv` and the `runnersupdata.csv` files. 
Save these in your R environment as `champs` and `runners` respectively. 

```{r}

```

```{r}

```

Take a look through these two datasets and see what columns and information is included. Use the `colnames()` function to print out what column names are in each. 

```{r}

```

There's a lot of acronyms that aren't clear what they mean. 

Go to https://www.kaggle.com/daverosenman/nba-finals-team-stats for more information on what these columns mean. 

### Fixing a typo

How many distinct teams are there? Used the `distinct()` and `select()` functions on the `champs` data frame and the `Team` variable to find out. 

```{r}

```

Notice anything strange about the distinct teams listed above? 

For example, `Warriorrs` and also `'Heat'`? 

Typos and errors in data happen all the time. Its part of your job as a data scientist to find this out and figure out how to handle these mistakes. 

Write code to replace all instances of `Warriorrs` with the correctly spelled `Warriors`. There are many ways you can do this. Use techniques we've covered and think about what makes sense to you to implement. 

Remember whatever strategy you use, once you have one figured out you'll need to make sure that the edited version of the `Team` column is re-saved back to the `champs` data frame over the old version of the `Team` column. 

```{r}

```

Does the `runners` data have this same problem? If so, use the code you figured out above to also fix this dataset. If not, then carry on your merry way. 

```{r}

```

### Fixing data types

Take a look at the columns `Win`, `Home`, and `Team` in the `champs` data frame. 

Find out what kind of data type each of these columns are using `str()` function. 

```{r}

```

```{r}

```


```{r}

```

Turn `Win` and `Home` into logical variables and, turn the `Team` column into a factor for both the `champs` and `runners` data frames. 

```{r}

```

### Summarize the data

Using the `dplyr::group_by()`, `dplyr::summarize()`, and `mean()` functions, find out the average total points (PTS) per game for a winning team by year. 

```{r}

```

Make a plot with `plot()` to see the average points over the years. Have there been higher total scores or any particular trends? 

```{r}

```

Find out this same question but for the runners up. 

```{r}

```

### Session Info 

```{r}
sessionInfo()
```
