---
title: "Getting Data"
author: "DataTrail"
date: "2022-09-22"
output: html_document
---

## Exercise 1: Testing URLs

1. Open up a new R script file (.R)
2. In this new script file, our goal is to create a function that takes a URL as input and returns a logical that tells us if the URL is real or not.
3. Let's use the httr package for this. Here's a useful tutorial: https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html
4. Test your resulting function with a URL that you know works as well as a fake URL that should fail. You will want your function to properly say TRUE when the URL is real and FALSE when the URL is fake.

```{r}
source(test_url.R)

# Should print out TRUE
test_url("https://www.datatrail.org/")

# Should print out FALSE
test_url("https://www.fakeurlauisdbviuabsduivbaiu.com")

# Should fail because 404
test_url("https://github.com/jhudsl/intro_to_r2")
```

## Exercise 2: Reading JSON files

1. Open up a new R Markdown file in RStudio.
2. Download this JSON file from GitHub: https://raw.githubusercontent.com/bahamas10/css-color-names/master/css-color-names.json
3. Use the jsonlite package to read in this file as a list. Here's a useful tutorial: https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html
4. Let's examine this resulting list. Subset this list to only contain the `peachpuff` color and its hex code.

```{r }
library(jsonlite)
```

```{r}
colors_json <- jsonlite::read_json("https://raw.githubusercontent.com/bahamas10/css-color-names/master/css-color-names.json")
```

```{r}
head(colors_json)
```

```{r}
colors_json$peachpuff
```

## Group Exercise 3: Let's chat about RSQLite

1. Here's this package's page: https://cran.r-project.org/web/packages/RSQLite/index.html 
2. Let's walk through together how we digest a package's documentation and know how to use it. 
3. RSQLite uses SQL -- a whole other language! So it's a bit of endeavor to use. But it can be a very handy tool if you have big databases.

- SQL commands: https://www.w3schools.com/sql/ 
- Practice SQL databases: https://learnsql.com/blog/free-online-datasets-to-practice-sql/

